---
title: "Git Cheat Sheet"
author: "Dobrin Petkov"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is a very brief list of the most used git commands with short explanations and simple examples.

# Getting started

After installing Git, it is good to customize your Git environment. The command to do it is:
```git
    git config
```

## Your identity

The first thing you should do is to set your user name and email address. This is used by Git in every commit:

```git
    git config --global user.name "Dobrin Petkov"
    git config --global user.email "dobrinpetkov@gmail.com"
```

## Your editor

Next you may want to set up your text editor. It is used whenever Git needs you to type in a message. If not set up, the default editor will be used:


```git
    git config --global core.editor emacs
```

## Your Settings

To check your settings just type:

```git
    git config --list
```

## Getting Help

From the shell:
```git 
    git help <verb>
    git <verb> help
    man git-<verb>
```

For example, if you want help on config:

```git 
    git help config
    git config help
    man git-config
```

# Typical Workflow - Single Branch

## Getting a Git Repository

There are two ways you can get a Git repository - initialize yourself or clone existing one.

### Initializen a Repository in an Existing Directory

To start tracking existing folder, navigate to it and then:

```git
    git init
```
This creates a new subdirectory called **.git** which contains all necessary repository files.

### Cloing an Existing Repository

```git
    git clone [url]
```
For example:
```git
    git clone https://github.com/dobrinpetkov/git_cheat_sheet
```
This creates directory `git_cheat_sheet` in your current folder, **.git** folder inside it, pulls down all the data for that repository, and checks out a working copy of the latest version. If you want to use custom folder name and not the default one, then:
```git
    git clone https://github.com/dobrinpetkov/git_cheat_sheet my_cheat_sheet
```

## Recording changes

Each file in your repository can be in one of the four states:

1. untracked
1. tracked
    - unmodified
    - modified
    - staged

A typical workflow is as follows:

- edit a file    (unmodified -> modified)
- stage the file (modified   -> staged)
- commit         (staged     -> unmodified)
- then the cycle repeats

### Checking the status of your files

```{r, engine = "bash", eval = T}
    git status
```
When you are comfortable enough with this command, you can use the short option:
```{r, engine = "bash", eval = T}
    git status -s
```

### Tracking New Files

```
    git add README
```

The `git add` takes a path name as input or a directory; **if it is a directory, it adds all the files in that directory recursively**.

### Ignoring files

If you want to ignore certain files, so they do not appear even in untracked ones, add them to the file `.gitignore`. You can use glob patterns. Those are simplified regular expressions:

- asterisk `*` matches zero or more characters
- `[abc]` matches any character inside the brackets (in this case a, b or c)
- a question mark `?` matches a single character
- `[0-9]` matches anything between 0 and 9
- `**` can be used to match nested directories. For example, **a/**/z** would match **a/z**, **a/b/z**, **a/b/c/z**, and s.o.

### Viewing Staged and Unstaged Changes

Let say you have changed a file, staged it, and then modified it again. 

```
    git diff
```
    
will show what is the difference between your modified version and your staged version. That is the difference between your working area and your staging area. If you want to see, what changes will be committed compared to your previous commit, run:

```
    git diff --staged (--cached) 
```
`cached` and `staged` are synonyms. Note that if you have only modified a file, and then run `git diff`, it will compare the versions in your working area and your previous commit.


### Committing Your Changes

```
    git commit
```
This will launch your editor of choice. There you can edit the commit message. To do it inline
```
    git commit -m "Your message goes here"
```
If you want, you might skip the staging are. 
```
    git commit -a 
```
Git will automatically stage every file, that has been tracked.

### Removing Files

To remove file from staging are, but keep it in the working tree (keep it on the hard, but do not track it with Git). This is useful if you have added files to the staging area which you did not want to (for example, forgetting to add something to `.gitingnore`).
```
    git rm --cached 
```


## Viewing the Commit History





